---
title: "Nb1_Data_validation_&_aggregation"
author: "Vivien Pichon"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)   # Data manipulation
library(reshape2)    # Data reshaping
library(knitr)       # Get good tables
library(kableExtra)  # Get good tables


knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE,
                      fig.align = 'center' )


## Create reporting function
describe_table <- function(df, name = deparse(substitute(df)), head = TRUE) {
  cat(sprintf("The table '%s' contains %d rows and %d columns\n", name, nrow(df), ncol(df)))
  if(head == TRUE){
    head(df) |>
      kable() |>
      kable_styling()
  }
}

# Simple rounded percentage function
turn_pct <- function(frac, total){
  round( 100*frac/total, 0)
}

```

# 1_Import tables

## 1.1 Non-filamentous

```{r}
## Import data sheets
# scores
tbl_score_nf <- read.csv( "data/0_nonfilamentous_database_cleaned.csv")

# strain characeristics
tbl_info <- read.csv( "data/0_strain_metadata.csv" )

## Filter and transform variables
tbl_score_nf <- tbl_score_nf |>
  mutate(patho_factor = factor(patho_factor,
                               levels = c("Dsol", "Pcar", "Asol",
                                          "Rsol", "Pinf", "Pult" ))
  )

## Keep key numbers
l_nonfil_collection_description <- list(
  ## Total number of non-filamentous
  total_nb_strains = n_distinct(tbl_score_nf$strain),
  ## Number of screened non-filamentous
  screened_strains_list = unique(tbl_score_nf$strain)
)


## Summarize table
describe_table(tbl_score_nf)

```


## 1.2 Filamentous


```{r ETL}

## Import data sheets
# scores
tbl_score_fil <- read.csv( "data/0_actino_score_database_cleaned.csv")

## Filter and transform variables
tbl_score_fil <- tbl_score_fil |> mutate(
  ## Turn pathogen to factor
  patho_factor = factor(patho_factor, 
                        levels = c("Dsol", "Pcar", "Asol",
                                   "Rsol", "Pinf", "Pult" ))
  )

## Summarize
describe_table(tbl_score_fil)


```


Import list of filamentous strains

```{r}

## list columns to keep
cols_to_keep <- c("strain", "Phylum", "Class", "Order", "Family", 
                  "Genus", "Medium", "Cultivar","compartment")

## Import actino list
tbl_actino_count <- read.csv( "data/0_fil_list_cleaned.csv") |>
  mutate(Phylum = PHYLUM, Class = CLASS, Order = ORDER, 
         Family = FAMILY, Genus = GENUS, Medium = MEDIUM,
         Cultivar = recode_values(SAMPLE,
                                  c("AS", "AR", "AL") ~ "Innovator",
                                  c("BS", "BR", "BL") ~ "Bintje",
                                  default = "error"
                                  ),
         compartment = recode_values(SAMPLE,
                                  c("AS", "BS") ~ "Soil",
                                  c("AR", "BR") ~ "Rhizosphere Soil",
                                  c("AL", "BL") ~"Leaf",
                                  default = "error"),
         .keep = "unused"
  ) |> 
  ## Select useful columns
  select(cols_to_keep, not_screened) |> 
  ## Replace NAs with 0s
  replace_na(list(not_screened = 0))

```

Place key numbers filamentous collection into a list

```{r}
l_filamentous_collection_description <- list(
  ## Total number of filamentous
  total_nb_strains = nrow(tbl_actino_count),
  ## Number of screened filamentous
  screened_strains_list = tbl_actino_count |> filter(not_screened == 0) |> pull(strain),
  ## number of failed 
  failed_strains_list = tbl_actino_count |> filter(not_screened == 1) |> pull(strain)
)

```


## 1.3 Compile collection descriptive numbers

```{r}

## Create table of all strains
table_all_strains <- tbl_score_nf |>
  ## Keep every distinct non-filamentous strain
  distinct(across(all_of(cols_to_keep))) |> 
  mutate(type = "nf") |> 
  bind_rows( 
    ## 
    tbl_actino_count |> 
      select(cols_to_keep) |> 
      mutate(type = "fil")
  )


## Create number-storing list
l_all_collection_description <- list(
  ## Total number of strains
  total_nb_strains = nrow(table_all_strains),
  ## per cultivar, number of isolates and genera
  isolates_per_cultivar = summarise(table_all_strains, n = n(), .by = Cultivar),
  genera_per_cultivar = summarise(table_all_strains, n = n_distinct(Genus), .by = Cultivar),
  ## per compartment, number of isolates and genera
  isolates_per_compartment = summarise(table_all_strains, n = n(), .by = compartment),
  genera_per_compartment = summarise(table_all_strains, n = n_distinct(Genus), .by = compartment),
  ## per growth medium, number of isolates and genera
  isolates_per_medium = summarise(table_all_strains, n = n(), .by = Medium),
  genera_per_medium = summarise(table_all_strains, n = n_distinct(Medium), .by = compartment),
  ## Per Phylum, number of isolates and percentage of the collection
  isolates_per_phylum = summarise(table_all_strains, n = n(), .by = Phylum),
  prct_per_phylum = summarise(table_all_strains, n = n(), .by = Phylum) |> mutate(n = turn_pct(n, nrow(table_all_strains))),
  ## Percentage of the collection per strain type (filamentous, non-fil)
  prct_per_type = summarise(table_all_strains, n = n(), .by = type) |> mutate(n = turn_pct(n, nrow(table_all_strains)))
  
)

```

# 2_Non-filamentous data wrangling
In this section we clean up the data by removing invalid rows, and aggregate it to end up with the table that we will use: A table categorizing the activity of each strain against each pathogen based on the median of the valid replicates.

## Validation

In this section, we: 

- remove the invalid replicates (score = 9, no sign of growth)
- remove the strains that are poorly suited for our protocol (strains with less than 10 valid replicates across all tests)
- remove the duels combination that are not reliable (only 1 valid replicate).
- remove the duels which have 2 replicates and a score difference between replicates > 1.

```{r replicate_growth_validation_nf}

### List useful columns
cols_to_keep <- c("strain", "patho_factor", 
                  "Phylum", "Class", "Order", "Family", "Genus", 
                  "Medium", "Cultivar","compartment")

## Remove invalid results
tbl_valid_nf = tbl_score_nf |>
  ## Return validation (0 is invalid, 1 is valid)
  mutate( validity = case_when(
    inhibition_score == 9 ~ 0,
    ctrl_growth == 0 & growth_score == 0 & inhibition_score == 0 ~ 0,
    is.na(ctrl_growth) & growth_score == 0 ~ 0,
    .default = 1)
    ) |>
  ## Remove invalid entries
  filter(validity == 1) |>
  ## select useful columnns
  select( c("inhibition_score",cols_to_keep) )

```

After cleaning, the table has the following properties.

```{r clean_strains_nf}
## Detect unsuited strains (<10 replicates)
list_bad_strains <- tbl_valid_nf |> 
  count(strain) |>
  filter(n < 10) |>
  pull(strain) |>
  unique()

## Remove unsuited strains
tbl_valid_nf <- tbl_valid_nf |>
  filter(!(strain %in% list_bad_strains))

## Report
cat(length(list_bad_strains), "strains were unsuited for our method.\n" )

## Detect unreliable combinations (less than 2 replicates)
list_bad_comb_nf <- tbl_valid_nf |> 
  count(strain, patho_factor) |>
  filter(n < 2) |>
  select(strain, patho_factor)

## Remove unsuited duels
tbl_valid_nf <- tbl_valid_nf |>
  anti_join(list_bad_comb_nf, by = c("strain", "patho_factor"))

## List duels with 2 replicates
list_2_duels_nf <- tbl_valid_nf |> 
  count(strain, patho_factor) |>
  filter(n==2) |> 
  select(-n)

## Find the duels for which the two replicates have a score difference >1
tbl_high_diff <- tbl_valid_nf |>
  inner_join(list_2_duels_nf, by = c("strain", "patho_factor")) |> 
  summarise(score_diff = first(inhibition_score)-last(inhibition_score),
            .by = c(strain, patho_factor)) |> 
  filter(score_diff > 1)

## Remove these duels
tbl_valid_nf <- tbl_valid_nf |>
  anti_join(tbl_high_diff, by = c("strain", "patho_factor"))

## Report
cat( nrow(list_bad_comb_nf) + nrow(tbl_high_diff), "duels were removed.\n" )

## Report
describe_table(tbl_valid_nf)


# Report data validation numbers
l_nonfil_data_validation <- list(
  ## Number of strains successfully screened
  nb_strains_screened = n_distinct(tbl_valid_nf$strain),
  ## Number of genera successfully screened
  nb_strains_genera = n_distinct(tbl_valid_nf$Genus),
  ## Number of strains with low number of replicates
  nb_invalid_less_than_10 = list_bad_strains |> length(),
  ## Number of duels with a single replicate
  nb_duel_with_one_replicate = list_bad_comb_nf |> nrow(),
  ## Number of strains with 2 replicates and a score difference >1
  nb_duels_highdiff = nrow(tbl_high_diff_fil),
  ## Add number of screened strains
  nb_strains_screened = tbl_valid_nf$strain |> n_distinct(), 
  ## Number of genera that the screened strains belong to
  nb_strains_genera = tbl_valid_nf$Genus |> n_distinct(),
  ## Total number of possible duels
  nb_possible_duels = l_nonfil_collection_description$total_nb_strains * 6,
  ## Actual number of duels
  nb_actual_duels = tbl_valid_nf |> count(strain, patho_factor) |> nrow(),
  ## Percentage of successfully assessed duels
  pct_duels_assessed = turn_pct(
    tbl_valid_nf |> count(strain, patho_factor) |> nrow(), 
    l_nonfil_collection_description$total_nb_strains * 6
    )
  )

```


## Aggregation

In this section we aggregate the data to deduce the activity level per strain per pathogen.
Firstly, the pathogen-specific scores are transformed to match the general scoring system (if they don't), then the scores are **scaled** to fit between 0 and 1. 
Then we categorize the activity level based to the **median** and the replicates:

- Median < 0.25 is **Inactive**.
- 0.25 <= Median < 0.5 has **Low** activity.
- 0.5 <= Median < 0.75 has **Medium** activity.
- 0.75 <= Median < 1 has **High** activity.
- Median == 1 has **Outstanding** activity.


```{r aggregation_nf}

tbl_agg_nf <- tbl_valid_nf |>
  ## Aggregate mean and median
  summarize(median = median(inhibition_score),
            .by = cols_to_keep
  ) |>
  ## Categorize median activity
  mutate( median_categ = case_when(
    median < 1 ~ "Inactive",
    median < 2 ~ "Low",
    median < 3 ~ "Medium",
    median < 4 ~ "High",
    median == 4 ~ "Outstanding",
    .default = "error") |>
      factor(levels = c("Outstanding", "High", "Medium", 
                        "Low", "Inactive", "Absent"))
  )

## Report
describe_table(tbl_agg_nf)

```


```{r save_nonfil_aggregated_data, eval=FALSE}

write.csv(tbl_agg_nf, "data/1_data_aggregated_non_filamentous",
          row.names = FALSE)

```

# 3_Filamentous Data Wrangling

In this section we do the same as previously, but with the filamentous strains.

## Validation

In the following section, we: 

- remove the invalid replicates (score = 9)
- remove the strains that are poorly suited for our protocol (strains with less than 10 valid replicates across all tests)
- remove the strain/pathogen combination that are not reliable (only 1 valid replicate).
- remove the duels which have 2 replicates and a score difference between replicates > 1.

```{r replicate_validation_fil}

## Remove invalid replicates (score =9)
tbl_valid_fil = tbl_score_fil |>
  filter( score != 9,
          !is.na(score)) |> 
  select(score, cols_to_keep)

## Remove strains with very low number of replicates (< 10 across all tests)
### Detect unsuited strains
list_bad_strains_fil <- tbl_valid_fil |> 
  count(strain) |>
  filter(n < 10) |>
  pull(strain) |>
  unique()

## Remove unsuited strains from dataset
tbl_valid_fil <- tbl_valid_fil |>
  filter(!(strain %in% list_bad_strains_fil))

## Report
cat( length(list_bad_strains_fil), "strains were unsuited for our method.\n" )

# Duels with a single replicate
## Detect
list_bad_comb_fil <- tbl_valid_fil |> 
  count(strain, patho_factor) |>
  filter(n < 2) |>
  select(strain, patho_factor)

## Remove these duels
tbl_valid_fil <- tbl_valid_fil |>
  anti_join(list_bad_comb_fil, by = c("strain", "patho_factor"))

# Duels with 2 replicates and a score difference >1 between them
## List duels with 2 replicates
list_2_duels_fil <- tbl_valid_fil |> 
  count(strain, patho_factor) |>
  filter(n==2) |> 
  select(-n)

## Find high difference duels 
tbl_high_diff_fil <- tbl_valid_fil |>
  inner_join(list_2_duels_fil, by = c("strain", "patho_factor")) |> 
  summarise(score_diff = first(score)-last(score),
            .by = c(strain, patho_factor)) |> 
  filter(score_diff > 1)

## Remove these duels
tbl_valid_fil <- tbl_valid_fil |>
  anti_join(tbl_high_diff_fil, by = c("strain", "patho_factor"))

## Report
cat( nrow(list_bad_comb_fil) + nrow(tbl_high_diff_fil), "combinations were removed.\n" )


# Report data validation numbers
l_filamentous_data_validation <- list(
  ## Number of strains with low number of replicates
  nb_invalid_less_than_10 = (list_bad_strains_fil |> length()) + 
    (tbl_actino_count |> filter(not_screened==1) |> nrow()),
  ## Number of duels with a single replicate
  nb_duel_with_one_replicate = list_bad_comb_fil |> nrow(),
  ## Number of strains with 2 replicates and a score difference >1
  nb_duels_highdiff = nrow(tbl_high_diff_fil),
  ## Add number of screened strains
  nb_strains_screened = tbl_valid_fil$strain |> n_distinct(), 
  ## Number of genera that the screened strains belong to
  nb_strains_genera = tbl_valid_fil$Genus |> n_distinct(),
  ## Total number of possible duels
  nb_possible_duels = l_filamentous_collection_description$filamentous_total * 6,
  ## Actual number of duels
  nb_actual_duels = tbl_valid_fil |> count(strain, patho_factor) |> nrow(),
  ## Percentage of successfully assessed duels
  pct_duels_assessed = turn_pct(
    (tbl_valid_fil |> count(strain, patho_factor) |> nrow()), 
    (l_filamentous_collection_description$filamentous_total * 6))
  )


nb_total_strains_not_screened <- (l_filamentous_collection_description$total_nb_strains - l_filamentous_data_validation$nb_strains_screened) + (l_nonfil_collection_description$total_nb_strains - l_nonfil_data_validation$nb_strains_screened)

```

## Aggregation

In this section we aggregate the data to deduce the activity level per strain per pathogen.
Firstly, the scores are **scaled** to fit between 0 and 1. 
Then we categorize the activity level based to the **median** and the replicates:

- Median < 0.25 is **Inactive**.
- 0.25 <= Median < 0.5 has **Low** activity.
- 0.5 <= Median < 0.75 has **Medium** activity.
- 0.75 <= Median < 1 has **High** activity.
- Median == 1 has **Outstanding** activity.

After analysis, the data looks as such.

```{r aggregation_fil}
## Aggregation
tbl_agg_fil <- tbl_valid_fil |>
  ## Aggregate median
  summarize(median = median(score),
            .by = cols_to_keep
            ) |>
      ## Categorize median activity
    mutate( median_categ = case_when(
      median < 1 ~ "Inactive",
      median < 2 ~ "Low",
      median < 3 ~ "Medium",
      median < 4 ~ "High",
      median == 4 ~ "Outstanding",
      .default = "error") |>
        factor(levels = c("Outstanding", "High", "Medium", 
                          "Low", "Inactive", "Absent"))
      )

## Report
describe_table(tbl_agg_fil)

```

```{r save_fil_aggregated_data, eval=FALSE}

write.csv(tbl_agg_fil, "data/1_data_aggregated_filamentous",
          row.names = FALSE)

```

## End

```{r get_SessionInfo}

sessionInfo()

``` 

