---
title: "Nb2_Figures_isolation_parameters_&_pathogen_sensitivity"
author: "Vivien Pichon"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)   # Data manipulation
library(reshape2)    # Data reshaping
library(knitr)       # Get good tables
library(kableExtra)  # Get good tables
library(RColorBrewer)# Color palette
library(patchwork)   # Pasting plots together
library(ggtext)      # Allows the use of the markdown text format in ggplots


knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE,
                      fig.align = 'center' )


## Create reporting function
describe_table <- function(df, name = deparse(substitute(df)), head = TRUE) {
  cat(sprintf("The table '%s' contains %d rows and %d columns\n", name, nrow(df), ncol(df)))
  if(head == TRUE){
    head(df) |>
      kable() |>
      kable_styling()
  }
}

# Simple rounded percentage function
turn_pct <- function(frac, total){
  round( 100*frac/total, 1)
}

```


# 1_Import data

```{r import_aggregated_tables}

## Import aggregated scores for non-filamentous strains
tbl_agg_nf <- read.csv("data/1_data_aggregated_non_filamentous") |> 
  ## Turn isolation factors and score category to factor
  mutate(Medium = factor(Medium,
                         levels = c("ARE", "TSA", "ActIA", "PDA")),
         compartment = factor(compartment,
                              levels = c("Leaf", "Root", "Rhizosphere soil")),
         Cultivar = factor(Cultivar,
                           levels = c("Bintje", "Innovator")),
         median_categ = factor(median_categ, 
                               levels = c("Outstanding", "High", "Medium", 
                                          "Low", "Inactive")),
         patho_factor = factor(patho_factor,
                               levels = c("Dsol", "Pcar", "Asol", "Rsol",
                                          "Pinf", "Pult"))
         )

## Import aggregated scores for non-filamentous strains
tbl_agg_fil <- read.csv("data/1_data_aggregated_filamentous") |> 
  ## Turn isolation factors and score category to factor
  mutate(Medium = factor(Medium,
                         levels = c("ARE", "TSA", "ActIA", "PDA")),
         compartment = factor(compartment,
                              levels = c("Leaf", "Root", "Rhizosphere soil")),
         Cultivar = factor(Cultivar,
                           levels = c("Bintje", "Innovator")),
         median_categ = factor(median_categ, 
                               levels = c("Outstanding", "High", "Medium", 
                                          "Low", "Inactive")),
         patho_factor = factor(patho_factor,
                               levels = c("Dsol", "Pcar", "Asol", "Rsol",
                                          "Pinf", "Pult"))
         )

## Merge filamentous and non-filamentous tables
tbl_agg_all <- tbl_agg_nf |> 
  mutate(type = "Non-Filamentous") |> 
  bind_rows( tbl_agg_fil |> mutate(type = "Filamentous")) |> 
  mutate(type = factor(type,
                       levels = c("Non-Filamentous", "Filamentous")))

```

# 2_Influence of isolation characteristics on maximum activity

We explore the influence the isolation factors on the strain activity.
For this, we will check if the distribution of the maximum activity level is influenced by any isolation factor: 

- **Growth medium** of isolation, 
- **cultivar** of origin,
- **compartment** of origin.

Firstly we create a table of maximum activity for each strain type.

The table looks like this:

```{r create_full_table}

## List factors to summarize by
summarize_by <- c("strain", "Medium", "Cultivar", "compartment")

## Create list with the 2 tables: filamentous, non-filamentous
l_max_activity <- list(tbl_fil = tbl_agg_fil |>
                         arrange(desc(median)) |> 
                         summarise(max_act = first(median_categ),
                                   .by = summarize_by),
                       tbl_nf = tbl_agg_nf |>
                         arrange(desc(median)) |> 
                         summarise(max_act = first(median_categ),
                                   .by = summarize_by))

## report
describe_table(l_max_activity$tbl_nf)

```

### *Figure 2*

We will represent the proportion of active strains for each factor, separately for the filamentous and non-filamentous strains.
We want the total number of strains below each bar.
Each subfigure will be for a different factor and will include the plots for the filamentous and non-filamentous strains.

Firstly we create the labels that contain the number of isolated from each value of each factor,

```{r, figure2_supporting_information}
## Attache list
attach(l_max_activity)

## Get the number of strains per value of each factor
list_strain_per_factor <- list(
  cultivar_nf = table(tbl_nf$Cultivar),
  cultivar_fil = table(tbl_fil$Cultivar),
  compartment_nf = table(tbl_nf$compartment),
  compartment_fil = table(tbl_fil$compartment),
  medium_nf = table(tbl_nf$Medium),
  medium_fil = table(tbl_fil$Medium)
)

## Attach list
attach(list_strain_per_factor)

## Arrange the numbers calculated above as small data frames that are usable by ggplot for labelling. 
list_factor_labels <- list(
  cultivar_nf = data.frame(
    Cultivar = names(cultivar_nf),
    label = as.character(cultivar_nf)
    ),
  cultivar_fil = data.frame(
    Cultivar = names(cultivar_fil),
    label = as.character(cultivar_fil)
    ),
  compartment_nf = data.frame(
    compartment = names(compartment_nf),
    label = as.character(compartment_nf)
    ),
  compartment_fil = data.frame(
    compartment = names(compartment_fil),
    label = as.character(compartment_fil)
  ),
  medium_nf = data.frame(
    medium = names(medium_nf),
    label = as.character(medium_nf)
  ),
  medium_fil = data.frame(
    medium = names(medium_fil),
    label = as.character(medium_fil)
  )
)

detach(list_strain_per_factor)
detach(l_max_activity)

```


Then we create the figure 2 by firstly plotting each subfigure, then arranging them together.

```{r figure_2_create}

attach(l_max_activity)
attach(list_factor_labels)


## Cultivar
# Plot 1: cultivar for non-filamentous
p1 <- ggplot(tbl_nf, aes(x = Cultivar, fill = max_act)) +
  geom_bar(position = "fill") +
  geom_richtext(
    data = cultivar_nf,
    aes(x = Cultivar, y = 1, label = paste("n =", label)),
    size = 2.5, 
    inherit.aes = FALSE, vjust = 0, fill = NA, label.color = NA,
    label.padding = unit(c(2, 0, 2, 0), "pt")
  ) +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "Non-Filamentous",
       fill = "Activity level",
       tag = "a") +
  theme_classic()+
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_text(),
    axis.text.x = element_blank(),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

# Plot 2: cultivar for filamentous
p2 <- ggplot(tbl_fil, aes(x = Cultivar, fill = max_act)) +
  geom_bar(position = "fill") +
  geom_richtext(
    data = cultivar_fil,
    aes(x = Cultivar, y = 1, label = paste("n =", label)),
    size = 2.5, 
    inherit.aes = FALSE, vjust = 0, fill = NA, label.color = NA,
    label.padding = unit(c(2, 0, 2, 0), "pt")
  ) +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "Filamentous",
       fill = "Activity level") +
  theme_classic()+
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_text(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

# Plot 3: compartment for non-filamentous
p3 <- ggplot(tbl_nf, aes(x = compartment, fill = max_act)) +
  geom_bar(position = "fill") +
  geom_richtext(
    data = compartment_nf,
    aes(x = compartment, y = 1, label = paste("n =", label)),
    size = 2.5, 
    inherit.aes = FALSE, vjust = 0, fill = NA, label.color = NA,
    label.padding = unit(c(2, 0, 2, 0), "pt")
  ) +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "",
       fill = "Activity level",
       tag = "b") +
  theme_classic()+
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

# Plot 4: compartment for filamentous
p4 <- ggplot(tbl_fil, aes(x = compartment, fill = max_act)) +
  geom_bar(position = "fill") +
  geom_richtext(
    data = compartment_fil,
    aes(x = compartment, y = 1, label = paste("n =", label)),
    size = 2.5, 
    inherit.aes = FALSE, vjust = 0, fill = NA, label.color = NA,
    label.padding = unit(c(2, 0, 2, 0), "pt")
  ) +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "",
       fill = "Activity level") +
  theme_classic()+
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

# Plot 5: medium for non-filamentous
p5 <- ggplot(tbl_nf, aes(x = Medium, fill = max_act)) +
  geom_bar(position = "fill") +
  geom_richtext(
    data = medium_nf,
    aes(x = medium, y = 1, label = paste("n =", label)),
    size = 2.3, 
    inherit.aes = FALSE, vjust = 0, fill = NA, label.color = NA,
    label.padding = unit(c(2, 0, 2, 0), "pt")
  ) +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "",
       fill = "Activity level",
       tag = "c") +
  theme_classic()+
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

# Plot 6: medium for filamentous
p6 <- ggplot(tbl_fil, aes(x = Medium, fill = max_act)) +
  geom_bar(position = "fill") +
  geom_richtext(
    data = medium_fil,
    aes(x = medium, y = 1, label = paste("n =", label)),
    size = 2.3, 
    inherit.aes = FALSE, vjust = 0, fill = NA, label.color = NA,
    label.padding = unit(c(2, 0, 2, 0), "pt")
  ) +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "",
       fill = "Activity level") +
  theme_classic()+
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")


figure_2 <- (p1|p3|p5)/(p2|p4|p6) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")

## Clean up
detach(l_max_activity)
detach(list_factor_labels)

rm(p1, p2, p3, p4, p5, p6)

figure_2

```

```{r, save_figure_2, include = FALSE}
ggsave("Figures/Figure_2.pdf",
       figure_2,
       height = 14,
       width = 21,
       units = "cm")

ggsave("Figures/Figure_2.png", 
       figure_2,
       height = 14,
       width = 21,
       units = "cm")

```


# 3_Influence of isolation parameters per pathogen

## *Figure S2*

Here, we create plots showing the influence of the isolation factor on the distribution of the activity, but this time for each pathogen separately.

```{r figure_S2_a}
## Set factor
fact <- "Cultivar"

figure_S2a <- ggplot(tbl_agg_all, aes(x = .data[[fact]], fill = median_categ )) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "",
       fill = "Activity level",
       tag = "a") +
  facet_grid(type ~ patho_factor) +
  theme_classic()+
  theme(
    strip.background = element_blank(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

figure_S2a

```

```{r figure_S2_b}
## Set factor
fact <- "compartment"

figure_S2b <- ggplot(tbl_agg_all, aes(x = .data[[fact]], fill = median_categ )) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "",
       fill = "Activity level",
       tag = "b") +
  facet_grid(type ~ patho_factor) +
  theme_classic()+
  theme(
    strip.background = element_blank(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

figure_S2b

```


```{r figure_S2_c}
## Set factor
fact <- "Medium"

figure_S2c <- ggplot(tbl_agg_all, aes(x = .data[[fact]], fill = median_categ )) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", 
       y = "",
       fill = "Activity level",
       tag = "c") +
  facet_grid(type ~ patho_factor) +
  theme_classic()+
  theme(
    strip.background = element_blank(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

figure_S2c

```

Assemble Figure S2

```{r assemble_figure_S2, fig.width=8 , fig.height=12}

figure_S2 <- (figure_S2a/figure_S2b/figure_S2c) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")

figure_S2

```

```{r save_figure_S2, include = FALSE}

## Save as pdf
ggsave("Figures/Supp_Figure_S2.pdf",
       figure_S2,
       height = 25,
       width = 17,
       units = "cm")

## Save as png
ggsave("Figures/Supp_Figure_S2.png",
       figure_S2,
       height = 25,
       width = 17,
       units = "cm")


```


# 4_Pathogen sensitivity to infection

## *Figure 3*

```{r, plot_figure, 3}

## Plot for non-filamentous strains
p1 <- ggplot(arrange(tbl_agg_nf, median_categ), 
             aes(x = patho_factor, y = 1, 
                 fill = median_categ)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual( values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  labs(tag = "a",
       x = "", y = "", 
       fill = "Activity") +
  theme_classic() +
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

## Plot for filamentous strains
p2 <-  ggplot(arrange(tbl_agg_fil, median_categ), 
       aes(x = patho_factor, y = 1, 
                    fill = median_categ)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual( values = c( brewer.pal(11, "RdGy")[2:5], "grey95")) +
  labs(tag = "b",
       x = "", y = "", 
       fill = "Activity") +
  theme_classic() +
  theme(
    axis.title.x = element_markdown(),
    panel.grid.major.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(),
    legend.title.position = "top") +
  coord_cartesian(clip = "off")

figure_3 <- (p1|p2) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")

rm(p1, p2)

figure_3

```

```{r, save_figure_3, include = FALSE}

ggsave("Figures/Figure_3.pdf",
       figure_3,
       height = 9,
       width = 18,
       units = "cm")

ggsave("Figures/Figure_3.png",
       figure_3,
       height = 9,
       width = 18,
       units = "cm")

```

# 5_Taxonomy of active strains
## *Figure S3*

```{r function_figure_S3}

func_s3_subfigure <- function(pathogen, tag) {
  
  ## Create vector of activity color
  activity_colors <- c(
    "Outstanding" = brewer.pal(11, "RdGy")[2],
    "High" = brewer.pal(11, "RdGy")[3],
    "Medium" = brewer.pal(11, "RdGy")[4],
    "Low" = brewer.pal(11, "RdGy")[5],
    "Inactive" = "grey90"
  )
  
  ## Create temporary table
  tbl_tmp <- tbl_agg_all |> filter(patho_factor == pathogen) |> arrange(median_categ)
  
  ## Create a data frame with the bar labels
  n_labels <- tbl_tmp %>%
    summarise(strain_nb = n_distinct(strain), .by = c(Phylum, Order))
  
  ## Create value plot
  fig <- ggplot(tbl_tmp, 
                aes(x = Order, y = 1, 
                    fill = median_categ)) +
    geom_bar(stat = "identity", position = "fill") +
    geom_text(data = n_labels,
              aes(x = Order, y = 1, label = paste0("n = ", strain_nb)),
              inherit.aes = FALSE,
              vjust = -0.5, size = 2) +
    scale_fill_manual( values = activity_colors) +
    labs(tag = tag,
         x = "", y = pathogen, 
         fill = "Activity") +
    facet_grid(. ~ Phylum, scales = "free_x", space = "free_x") +
    theme_classic() +
    theme(
      axis.title.x = element_markdown(),
      panel.grid.major.x = element_blank(),
      axis.text.y = element_blank(),
      axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1, size = 4),
      legend.title = element_text(),
      legend.title.position = "top",
      legend.position = "none",
      strip.background = element_blank(),
      strip.text = element_text(margin = margin(b = 10), size = 6),
      panel.spacing = unit(0.15, "lines")) +
    coord_cartesian(clip = "off")
  
  return(fig)
}

  
```


```{r create_figure_S3, fig.height=12, fig.width=8}

s3a <- func_s3_subfigure("Dsol", "a")
s3b <- func_s3_subfigure("Pcar", "b")
s3c <- func_s3_subfigure("Asol", "c")
s3d <- func_s3_subfigure("Rsol", "d")
s3e <- func_s3_subfigure("Pinf", "e") + theme(legend.position = "right")
s3f <- func_s3_subfigure("Pult", "f")


figure_S3 <- (s3a/s3b/s3c/s3d/s3e/s3f) + 
  plot_layout(guides = "collect")

rm(s3a, s3b, s3c, s3d, s3e, s3f)

figure_S3


```


```{r, save_figure_S3, include = FALSE}

ggsave("Figures/Supp_Figure_S3.pdf",
       figure_S3,
       height = 30,
       width = 18,
       units = "cm")

ggsave("Figures/Supp_Figure_S3.png",
       figure_S3,
       height = 30,
       width = 18,
       units = "cm")

```


# 6_Key data collection

## Isolation parameters


```{r}
attach(l_max_activity)

## List non-filamentous strains
l_isolation_parameters_nf <- list(
  ## Proportion of active strains 
  prop_active_strains = turn_pct( (tbl_nf$max_act!="Inactive") |> sum(), tbl_nf |> nrow()),
  ## Proportion active strains per cultivar
  prop_active_strains_per_cultivar = tbl_nf |> 
    summarise(pct_active = turn_pct(sum(max_act!="Inactive"), n()),
              .by = Cultivar),
  ## Proportion active strains per compartment
  prop_active_strains_per_compartment = tbl_nf |> 
    summarise(pct_active = turn_pct(sum(max_act!="Inactive"), n()),
              .by = compartment),
  ## Proportion active strains per growth medium
  prop_active_strains_per_medium = tbl_nf |> 
    summarise(pct_active = turn_pct(sum(max_act!="Inactive"), n()),
              .by = Medium)
)


### List for filamentous strains
l_isolation_parameters_fil <- list(
  ## Proportion of active strains 
  prop_active_strains = turn_pct( (tbl_fil$max_act!="Inactive") |> sum(), tbl_fil |> nrow()),
  ## Proportion active strains per cultivar
  prop_active_strains_per_cultivar = tbl_fil |> 
    summarise(pct_active = turn_pct(sum(max_act!="Inactive"), n()),
              .by = Cultivar),
  ## Proportion active strains per compartment
  prop_active_strains_per_compartment = tbl_fil |> 
    summarise(pct_active = turn_pct(sum(max_act!="Inactive"), n()),
              .by = compartment),
  ## Proportion active strains per growth medium
  prop_active_strains_per_medium = tbl_fil |> 
    summarise( pct_active = turn_pct( sum( max_act != "Inactive"), n()),
              .by = Medium)
)


detach(l_max_activity)

```

## Pathogen sensitivity

```{r}

## List non-filamentous
l_patho_sensitivity_nf <- list(
  ## Number and percentage of active strains per pathogen
  active_strains_per_pathogen = tbl_agg_nf |> 
    summarise(count_active = sum( median_categ != "Inactive" ),
              pct_active = turn_pct( sum( median_categ != "Inactive" ), n() ),
              .by = patho_factor ) |> 
    arrange(patho_factor),
  
  ## Number and percentage of strains per pathogen for each activity level
  activity_levels_per_pathogen = tbl_agg_nf |> 
    mutate(n_strain = n_distinct(strain),
           .by = patho_factor) |> 
    summarise(count = n(),
              pct = turn_pct( n(), first(n_strain)),
              .by = c(patho_factor, median_categ)) |> 
    arrange(patho_factor, median_categ)
)


## List filamentous
l_patho_sensitivity_fil <- list(
  ## Number and percentage of active strains per pathogen
  active_strains_per_pathogen = tbl_agg_fil |> 
    summarise(count_active = sum(median_categ!="Inactive"),
              pct_active = turn_pct(sum(median_categ!="Inactive"), n()),
              .by = patho_factor) |> 
    arrange(patho_factor),
  
  ## Number and percentage of strains per pathogen for each activity level
  activity_levels_per_pathogen = tbl_agg_fil |> 
    mutate(n_strain = n_distinct(strain),
           .by = patho_factor) |> 
    summarise(count = n(),
              pct = turn_pct( n(), first(n_strain)),
              .by = c(patho_factor, median_categ))|> 
    arrange(patho_factor, median_categ)
  
)


```


# End

```{r get SessionInfo}

sessionInfo()

```


