---
title: "Nb3_phylogenetic_trees"
author: "Vivien Pichon"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)   # Data manipulation
library(reshape2)    # Data reshaping
library(knitr)       # Get good tables
library(kableExtra)  # Get good tables
library(RColorBrewer)# Color palette
library(patchwork)   # Pasting plots together
library(ggtext)      # Allows the use of the markdown text format in ggplots
library(ggtree)      # Build the phylogenetic trees
library(ggtreeExtra) # Extra functionalities for ggtree
library(pheatmap)    # Good heatmap package
library(ggnewscale)  # Create new scales for ggplot
library(geiger)      # Rescale the full phylogenetic tree
library(cowplot)     # Make composite plots

knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE,
                      fig.align = 'center' )


## Create reporting function
describe_table <- function(df, name = deparse(substitute(df)), head = TRUE) {
  cat(sprintf("The table '%s' contains %d rows and %d columns\n", name, nrow(df), ncol(df)))
  if(head == TRUE){
    head(df) |>
      kable() |>
      kable_styling()
  }
}

# Simple rounded percentage function
turn_pct <- function(frac, total){
  round( 100*frac/total, 1)
}

```

In this notebook we create phylogenetic trees of the collection, annotated with a heatmap that contains the activity level of each strain against each pathogen.
Two figures, each containing two phylogenetic trees (non-filamentous, filamentous) will be created. 
The first one contains all the strains but no strain labels (Figure 4). 
The second one contains only the active strains, and has strain labels (Figure S4).

# 0_Import data

```{r import_aggregated_tables}

## Import aggregated scores for non-filamentous strains
tbl_agg_nf <- read.csv("data/1_data_aggregated_non_filamentous.csv") |> 
  ## Turn isolation factors and score category to factor
  mutate(Medium = factor(Medium,
                         levels = c("ARE", "TSA", "ActIA", "PDA")),
         compartment = factor(compartment,
                              levels = c("Leaf", "Root", "Rhizosphere soil")),
         Cultivar = factor(Cultivar,
                           levels = c("Bintje", "Innovator")),
         median_categ = factor(median_categ, 
                               levels = c("Outstanding", "High", "Medium", 
                                          "Low", "Inactive")),
         patho_factor = factor(patho_factor,
                               levels = c("Dsol", "Pcar", "Asol", "Rsol",
                                          "Pinf", "Pult"))
         )

## Import aggregated scores for non-filamentous strains
tbl_agg_fil <- read.csv("data/1_data_aggregated_filamentous.csv") |> 
  ## Turn isolation factors and score category to factor
  mutate(Medium = factor(Medium,
                         levels = c("ARE", "TSA", "ActIA", "PDA")),
         compartment = factor(compartment,
                              levels = c("Leaf", "Root", "Rhizosphere soil")),
         Cultivar = factor(Cultivar,
                           levels = c("Bintje", "Innovator")),
         median_categ = factor(median_categ, 
                               levels = c("Outstanding", "High", "Medium", 
                                          "Low", "Inactive")),
         patho_factor = factor(patho_factor,
                               levels = c("Dsol", "Pcar", "Asol", "Rsol",
                                          "Pinf", "Pult"))
         )


## list columns to keep
cols_to_keep <- c("strain", "Phylum", "Class", "Order", "Family", 
                  "Genus", "Medium", "Cultivar","compartment")

## Import full collection phylogenetic tree
tree_full <- read.tree( "data/0_full_tree.nwk")

## Import filamentous phylogenetic tree
tree_fil <- read.tree( "data/0_filamentous_tree.nwk")


```

# 1_Big trees: Supporting objects

## 1.1_Non-Filamentous

### Info table

```{r create_info_tab_big_tree_nf}

## Create list of objects for the tree
list_tree_big_nf <- list(all_strains = tbl_agg_nf$strain |> unique() |> sort())

## attach the list
attach(list_tree_big_nf)

## Create info table for small tree
# Import full table
list_tree_big_nf <- list_tree_big_nf |> 
  list_assign(info = tbl_agg_nf |>
                # Keep only one row of each strain
                distinct(across(all_of(cols_to_keep))) |>
                # Select relevant phylogeny columns
                select(cols_to_keep) |>
                # Turn strain column to row names
                column_to_rownames("strain")
  )

##Report
describe_table(list_tree_big_nf$info)

detach(list_tree_big_nf)

```

### Heatmap data frame

```{r create_heatmaps_big_tree_nf}

## Initiate results heatmap as a data frame
mat_active_tmp <- tbl_agg_nf |>
  ## code activity level into numbers
  mutate(act_code = recode_values(median_categ,
                                  "Inactive" ~ 0,
                                  "Low" ~ 1,
                                  "Medium" ~ 2,
                                  "High" ~ 3,
                                  "Outstanding" ~ 4)) |> 
  ## Pivot table into matrix-type
  pivot_wider(names_from = patho_factor,
              id_cols = strain,
              values_from = act_code) |> 
  ## Create row names
  column_to_rownames("strain") |> 
  select("Dsol", "Pcar", "Asol", "Rsol", "Pinf", "Pult")

## Add spaces in the column names for proper alignment on the circular phylogenetic tree
colnames(mat_active_tmp) <- c("Dsol", "  Pcar", 
                              "   Asol", "   Rsol", 
                              "     Pinf", "       Pult")

# # Convert all columns to factor, keeping consistent levels 0-4
mat_active_tmp[] <- lapply(mat_active_tmp, 
                           function(x) factor(x, levels = 0:4))

## Add to the list
list_tree_big_nf <- list_tree_big_nf |> 
  list_assign(results_hm = mat_active_tmp)

## Clean up temporary object
rm(mat_active_tmp)

## Report
describe_table(list_tree_big_nf$results_hm)

```

### Taxalist

```{r create_taxalist_big_tree_nf}

## attach the list we work with
attach(list_tree_big_nf)

## Create empty taxalist object (branch coloring the trees)
# Choose rank (Phylum, Order, Genus)
rank = "Order"

# Create empty object
taxalist_tmp <- NULL


# Fill objects as a list with vectors, all named "i" 
for (i in unique( info[ , rank] )) {
  taxalist_tmp <- c(taxalist_tmp, 
                    list( i = rownames(info)[info[ , rank] == i]))
}

# Rename the vectors of the list after their taxon
names(taxalist_tmp) <- unique(info[ , rank])

## List taxa with no active strain
other_taxa <- c( names(taxalist_tmp)[!(names(taxalist_tmp) %in%
                                     (tbl_agg_nf |> filter(median != 0) |> pull(rank) |> unique() )
                                      )],
                 "Enterobacterales", "Brevibacillales")

# Merge the contents of the other_taxa elements into one vector
taxalist_tmp[["Other"]] <- unlist(taxalist_tmp[names(taxalist_tmp) %in% other_taxa])

# Remove the original elements
taxalist_tmp <- taxalist_tmp[!names(taxalist_tmp) %in% other_taxa]

## Add the taxa list to the tree list
list_tree_big_nf <- list_tree_big_nf |> 
  list_assign(taxalist = taxalist_tmp)

## Clean up temporary items
rm(taxalist_tmp, other_taxa, rank, i)

detach(list_tree_big_nf)

```

### Format tree

```{r format_tree_big_nf}

## Attach list
attach(list_tree_big_nf)

## Prune and recalculate the tree with the used strains only
tree_tmp <- ape::keep.tip(
  tree_full, 
  rownames(results_hm)) |>
  ## Root the tree on AR046
  ape::root("AR046") |> 
  ## Rescale tree
  rescale( model = "delta", delta = 0.7)

## Add the tree to the list
list_tree_big_nf <- list_tree_big_nf |> 
  list_assign(tree = tree_tmp)


## Clean up temporary item
rm(tree_tmp)
detach(list_tree_big_nf)

```

## 1.2_Filamentous

### Info table

```{r create_info_tab_big_tree_fil}

## Create list of objects for the tree
list_tree_big_fil <- list(all_strains = tbl_agg_fil$strain |> unique() |> sort())

## attach the list
attach(list_tree_big_fil)

## Create info table for small tree
# Import full table
list_tree_big_fil <- list_tree_big_fil |> 
  list_assign(info = tbl_agg_fil |>
                # Keep only one row of each strain
                distinct(across(all_of(cols_to_keep))) |>
                # Select relevant phylogeny columns
                select(cols_to_keep) |>
                # Turn strain column to row names
                column_to_rownames("strain")
  )

##Report
describe_table(list_tree_big_fil$info)

detach(list_tree_big_fil)

```

### Heatmap data frame

```{r create_heatmaps_big_tree_fil}

## Initiate results heatmap as a data frame
mat_active_tmp <- tbl_agg_fil |>
  ## code activity level into numbers
  mutate(act_code = recode_values(median_categ,
                                  "Inactive" ~ 0,
                                  "Low" ~ 1,
                                  "Medium" ~ 2,
                                  "High" ~ 3,
                                  "Outstanding" ~ 4)) |> 
  ## Pivot table into matrix-type
  pivot_wider(names_from = patho_factor,
              id_cols = strain,
              values_from = act_code) |> 
  ## Create row names
  column_to_rownames("strain") |> 
  select("Dsol", "Pcar", "Asol", "Rsol", "Pinf", "Pult")

## Add spaces in the column names for proper alignment on the circular phylogenetic tree
colnames(mat_active_tmp) <- c("Dsol", "  Pcar", 
                              "   Asol", "   Rsol", 
                              "     Pinf", "       Pult")

# # Convert all columns to factor, keeping consistent levels 0-4
mat_active_tmp[] <- lapply(mat_active_tmp, 
                           function(x) factor(x, levels = 0:4))

## Add to the list
list_tree_big_fil <- list_tree_big_fil |> 
  list_assign(results_hm = mat_active_tmp)

## Clean up temporary object
rm(mat_active_tmp)

## Report
describe_table(list_tree_big_fil$results_hm)

```

### Taxalist

```{r create_taxalist_big_tree_fil}

## attach the list we work with
attach(list_tree_big_fil)

## Create empty taxalist object (branch coloring the trees)
# Choose rank (Phylum, Order, Genus)
rank = "Order"

# Create empty object
taxalist_tmp <- NULL


# Fill objects as a list with vectors, all named "i" 
for (i in unique( info[ , rank] )) {
  taxalist_tmp <- c(taxalist_tmp, 
                    list( i = rownames(info)[info[ , rank] == i]))
}

# Rename the vectors of the list after their taxon
names(taxalist_tmp) <- unique(info[ , rank])


## Add the taxa list to the tree list
list_tree_big_fil <- list_tree_big_fil |> 
  list_assign(taxalist = taxalist_tmp)

## Clean up temporary items
rm(taxalist_tmp, other_taxa, rank, i)

detach(list_tree_big_fil)

```

### Format tree

```{r format_tree_big_fil}

## Attach list
attach(list_tree_big_fil)

## Prune and recalculate the tree with the used strains only
tree_tmp <- ape::keep.tip(
  tree_fil, 
  rownames(results_hm)) |>
  ## Root the tree on AR046
  ape::root("BR162", resolve.root = TRUE) |> 
  ## Rescale tree
  rescale( model = "delta", delta = 0.7)

## Add the tree to the list
list_tree_big_fil <- list_tree_big_fil |> 
  list_assign(tree = tree_tmp)


## Clean up temporary item
rm(tree_tmp)
detach(list_tree_big_fil)

```

# 2_Big trees: Construction

## 2.1_Non-filamentous *Figure4a*

```{r plot_big_tree_nf, fig.align='left', fig.height=15, fig.width=19}

## Attach list
attach(list_tree_big_nf)

p <- ggtree(tree,
            ladderize = TRUE,
            linewidth = .7,
            layout = "fan",
            open.angle = 10) %<+% info +
  # Add metadata and tiplabs
  geom_tiplab(size = 0, 
              align = TRUE, 
              linesize = 0.2,
              linetype = "solid",
              alpha = .3,
              show.legend = FALSE) +
  labs(tag = "a")+
  theme(legend.title = element_text(size = 10),
        plot.title = element_text(margin = margin(b = 2, unit = "mm"),
                                  vjust = -10),
        legend.text = element_text(size = 9),
        plot.margin = margin(-20, -20, 0, -20, "pt"),
        plot.tag.position = c(0.05, 0.90))

## Color the branches
name_order <- c("Corynebacteriales",  "Micrococcales", 
                "Bacillales", 
                "Caulobacterales", "Rhizobiales", 
                "Sphingobacteriales", "Pseudomonadales", "Burkholderiales", "Xanthomonadales",
                "Other")

q = groupOTU(p, taxalist, "Order") +
  aes(color = Order) +
  scale_color_manual( breaks = names(taxalist)[
    match(name_order, names(taxalist))],
    values = c(
      "limegreen", "olivedrab3",  
      "steelblue1", 
      "purple1", "royalblue2",
      "goldenrod1",
      "violetred", "hotpink1","deeppink2",
      "grey", "black" ))


## Add categorical results heatmap to base tree
figure_4a <- gheatmap(q,
                      results_hm,
                      offset = 0,
                      width = 0.5,
                      color = "white",
                      colnames_angle = 55,
                      font.size = 1.6,
                      colnames_offset_y = 369,
                      colnames_offset_x = -0.03
) +
  scale_fill_manual(
    values = c("0" = "grey95",
               "1" = brewer.pal(11, "RdGy")[5],
               "2" = brewer.pal(11, "RdGy")[4],
               "3" = brewer.pal(11, "RdGy")[3],
               "4" = brewer.pal(11, "RdGy")[2]),
    labels = c("Inactive",
               "Low",
               "Medium",
               "High",
               "Outstanding"),
    na.value = "white",
    name = "Activity") + 
  hexpand(0.1, direction = 1)


## Clean up
rm(p, q, name_order)
detach(list_tree_big_nf)

print(figure_4a)

```

## 2.2_Filamentous *Figure4b*

```{r plot_fig4b, fig.align='left', fig.height=15, fig.width=19}

## Attach list
attach(list_tree_big_fil)

p <- ggtree(tree,
            ladderize = TRUE,
            linewidth = .7,
            layout = "fan",
            open.angle = 10) %<+% info +
  # Add metadata and tiplabs
  geom_tiplab(size = 0,
              align = TRUE,
              linesize = .2,
              linetype = "solid",
              alpha = 1,
              show.legend = FALSE) +
  labs(tag = "b")+
  theme(legend.title = element_text(size = 10),
        plot.title = element_text(margin = margin(b = 2, unit = "mm"),
                                  vjust = -10),
        legend.text = element_text(size = 9),
        plot.margin = margin(-20, -20, -20, -20, "pt"),
        plot.tag.position = c(0.05, 0.90)) + 
  hexpand(0.1, direction = 1)

q <- groupOTU(p, taxalist, "Genus") +
  aes(color = Genus) +
  scale_color_manual( breaks = names(taxalist)[match(c("Streptomycetales",
                                                       "Propionibacteriales",
                                                       "Corynebacteriales"), names(taxalist))],
                      values = c("limegreen", "steelblue1",
                                 "goldenrod1", "grey", "black"))


## Add categorical results heatmap to base tree
figure_4b <- gheatmap(q,
              results_hm,
              offset = 0.0,
              width = 0.5,
              color = "white",
              colnames_angle = 50,
              font.size = 1.6,
              colnames_offset_y = 205.5,
              colnames_offset_x = -0.006
) +
  scale_fill_manual(
    values = c("0" = "grey95",
               "1" = brewer.pal(11, "RdGy")[5],
               "2" = brewer.pal(11, "RdGy")[4],
               "3" = brewer.pal(11, "RdGy")[3],
               "4" = brewer.pal(11, "RdGy")[2]),
    labels = c("Inactive",
               "Low",
               "Medium",
               "High",
               "Outstanding"),
    na.value = "white",
    name = "Activity") + 
  hexpand(0.1, direction = 1)


## Clean up
rm(p, q, name_order)
detach(list_tree_big_fil)

print(figure_4b)

```


## Assemble *Figure 4*

```{r, create_figure_4}

## Extract each legend individually, suppressing the other guide when doing so
legend_color1 <- get_legend(figure_4a + theme(legend.position = "right") + guides(fill = "none"))
legend_fill   <- get_legend(figure_4a + theme(legend.position = "right") + guides(color = "none"))
legend_color2 <- get_legend(figure_4b + theme(legend.position = "right") + guides(fill = "none"))

## Remove all legends from the tree plots themselves
figure_4a <- figure_4a + theme(legend.position = "none")
figure_4b <- figure_4b + theme(legend.position = "none")

## Stack the three legends in the desired order
legends_stacked <- wrap_plots(
  legend_color1, legend_fill, legend_color2,
  ncol = 1
)

figure_pre4 <- (wrap_elements(full = figure_4a) / wrap_elements(full = figure_4b)) +
  plot_layout(heights = c(1.1, 1))

figure_4 <- (figure_pre4 | legends_stacked) +
  plot_layout(widths = c(1, 0.35))

figure_4

``` 

```{r save_figure_4, include = FALSE}

ggsave("Figures/Figure_4.pdf",
       figure_4,
       height = 20,
       width = 17,
       units = "cm")

ggsave("Figures/Figure_4.png",
       figure_4,
       height = 20,
       width = 17,
       units = "cm",
       dpi = 600)

```


## 2.3_Figure 4b for a poster

```{r plot_fig4b_fil_poster}

attach(list_tree_big_fil)

#Add a scale factor for all scalable elements
scale_factor <- 2

p <- ggtree(tree,
            ladderize = TRUE,
            linewidth = 1 * scale_factor * 0.4,
            layout = "fan",
            open.angle = 10) %<+% info +
  geom_tiplab(size = 0,
              align = TRUE,
              linesize = .2 * scale_factor * 0.4,
              linetype = "solid",
              alpha = 1,
              show.legend = FALSE) + 
  theme(legend.title = element_text(size = 10 * scale_factor),
        legend.text   = element_text(size = 9 * scale_factor),
        plot.tag      = element_text(size = 14 * scale_factor, face = "bold"),
        plot.title    = element_text(margin = margin(b = 2, unit = "mm"),
                                      vjust = -10),
        plot.margin   = margin(0, 0, 0, 0, "pt"),
        legend.key.size = unit(scale_factor * 4, "mm"),
        plot.tag.position = c(0.05, 0.90)) +
  hexpand(0.1, direction = 1)

q <- groupOTU(p, taxalist, "Genus") +
  aes(color = Genus) +
  scale_color_manual(breaks = names(taxalist)[match(c("Streptomyces", "Nocardioides", "Nocardia"), names(taxalist))],
                      values = c("limegreen", "steelblue1",
                                 "goldenrod1", "grey", "black"))

## Add categorical results heatmap to base tree
figure_4b_poster <- gheatmap(q,
              results_hm,
              offset = 0.0,
              width = 0.5,
              color = "white",
              colnames_angle = 50,
              font.size = 1.6 * scale_factor,
              colnames_offset_y = 206,
              colnames_offset_x = -0.006
) +
  scale_fill_manual(
    values = c("0" = "grey95",
               "1" = brewer.pal(11, "RdGy")[5],
               "2" = brewer.pal(11, "RdGy")[4],
               "3" = brewer.pal(11, "RdGy")[3],
               "4" = brewer.pal(11, "RdGy")[2]),
    labels = c("Inactive", "Low", "Medium", "High", "Outstanding"),
    na.value = "white",
    name = "Activity") +
  hexpand(0.1, direction = 1) +
  theme(legend.title = element_text(size = 10 * scale_factor),
        legend.text   = element_text(size = 9 * scale_factor))

rm(p, q)
detach(list_tree_big_fil)


print(figure_4b_poster)

```

```{r save_figure_4b_poster, include = FALSE}

ggsave("Figures/Figure_4b_poster.pdf",
       figure_4b_poster,
       height = 25,
       width = 30,
       units = "cm",
       dpi = 600)

ggsave("Figures/Figure_4b_poster.png",
       figure_4b_poster,
       height = 25,
       width = 35,
       units = "cm",
       dpi = 600)

```

# 3_Small trees: Supporting objects

The small trees are the labelled phylogenetic trees with only the active strains.
Like for the big trees, we start with creating the supporting files, and then construct the trees.

## 3.1_Non-filamentous

```{r small_tree_nf_supporting_objects}

attach(list_tree_big_nf)

list_active_strains = tbl_agg_nf |> filter(median_categ != "Inactive") |> pull(strain) |> unique() |> sort()

list_tree_small_nf <- list(
  ## Subset info
  info = info[list_active_strains,],
  results_hm = results_hm[list_active_strains,],
  tree =  
    ## Keep only active strains
    ape::keep.tip(tree_full, 
                  list_active_strains) |>
    ## Root the tree on AS109
    ape::root("AS109") |>
    ## Rescale tree
    rescale( model = "delta", delta = 0.7),
  taxalist = taxalist |> 
    lapply(function(x) x[x %in% list_active_strains]) %>% ## Use %>% because it supports the "." placeholder
    Filter(function(x) length(x) > 0, .) 
  
)

rm(list_active_strains)

detach(list_tree_big_nf)

```

## 3.2_Filamentous

```{r small_tree_fil_supporting_objects}

attach(list_tree_big_fil)

list_active_strains = tbl_agg_fil |> filter(median_categ != "Inactive") |> pull(strain) |> unique() |> sort()

list_tree_small_fil <- list(
  ## Subset info
  info = info[list_active_strains,],
  results_hm = results_hm[list_active_strains,],
  tree =  
    ## Keep only active strains
    ape::keep.tip(tree_fil, 
                  list_active_strains) |>
    ## Root the tree on BR162
    ape::root("BR162") |>
    ## Rescale tree
    rescale( model = "delta", delta = 0.7),
  taxalist = taxalist |> 
    lapply(function(x) x[x %in% list_active_strains]) %>% ## Use %>% because it supports the "." placeholder
    Filter(function(x) length(x) > 0, .) 
  
)

rm(list_active_strains)

detach(list_tree_big_fil)

```

# 4_Small trees: Construct
 
## Non-filamentous *Figure S4a*

```{r plot_small_tree_nf, fig.align='left', fig.height=15, fig.width=19}

## Attach list
attach(list_tree_small_nf)

p <- ggtree(tree,
            ladderize = TRUE,
            linewidth = .7,
            layout = "fan",
            open.angle = 10) %<+% info +
  # Add metadata and tiplabs
  geom_tiplab(size = 1.8, 
              align = TRUE, 
              linesize = 0.2,
              linetype = "solid",
              alpha = 1,
              show.legend = FALSE) +
  labs(tag = "a")+
  theme(legend.title = element_text(size = 10),
        plot.title = element_text(margin = margin(b = 2, unit = "mm"),
                                  vjust = -10),
        legend.text = element_text(size = 9),
        plot.margin = margin(-20, -20, 0, -20, "pt"),
        plot.tag.position = c(0.05, 0.90))

## Order of the taxa in the legend
name_order <- c("Corynebacteriales",  "Micrococcales",
                "Bacillales",
                "Caulobacterales", "Rhizobiales",
                "Sphingobacteriales", "Pseudomonadales", "Burkholderiales", "Xanthomonadales" )

q = groupOTU(p, taxalist, "Order") +
  aes(color = Order) +
  scale_color_manual( breaks = names(taxalist)[
    match(name_order, names(taxalist))],
    values = c(
      Corynebacteriales = "limegreen", 
      Micrococcales = "olivedrab3",  
      Bacillales = "steelblue1", 
      Caulobacterales = "purple1",
      Rhizobiales = "royalblue2",
      Sphingobacteriales = "goldenrod1",
      Pseudomonadales = "violetred", 
      Burkholderiales = "hotpink1",
      Xanthomonadales = "deeppink2" ))


## Add categorical results heatmap to base tree
figure_S4a <- gheatmap(q,
                      results_hm,
                      offset = .22,
                      width = 0.5,
                      color = "white",
                      colnames_angle = 55,
                      font.size = 1.6,
                      colnames_offset_y = 95,
                      colnames_offset_x = -0.03
) +
  scale_fill_manual(
    values = c("0" = "grey95",
               "1" = brewer.pal(11, "RdGy")[5],
               "2" = brewer.pal(11, "RdGy")[4],
               "3" = brewer.pal(11, "RdGy")[3],
               "4" = brewer.pal(11, "RdGy")[2]),
    labels = c("Inactive",
               "Low",
               "Medium",
               "High",
               "Outstanding"),
    na.value = "white",
    name = "Activity")


## Clean up
rm(p, q, name_order)
detach(list_tree_small_nf)

print(figure_S4a)

```

## Filamentous *Figure S4b*

```{r plot_small_tree_fil, fig.align='left', fig.height=10, fig.width=14}

## Attach list
attach(list_tree_small_fil)

p <- ggtree(tree,
            ladderize = TRUE,
            linewidth = .7,
            layout = "fan",
            open.angle = 10) %<+% info +
  # Add metadata and tiplabs
  geom_tiplab(size = 1.5, 
              align = TRUE, 
              linesize = 0.2,
              linetype = "solid",
              alpha = 1,
              show.legend = FALSE) +
  labs(tag = "b")+
  theme(legend.title = element_text(size = 10),
        plot.title = element_text(margin = margin(b = 2, unit = "mm"),
                                  vjust = -10),
        legend.text = element_text(size = 9),
        plot.margin = margin(-20, -20, 0, -20, "pt"),
        plot.tag.position = c(0.05, 0.90))


q <- groupOTU(p, taxalist, "Genus") +
  aes(color = Genus) +
  scale_color_manual( breaks = names(taxalist)[match(c(
    "Streptomycetales", "Propionibacteriales", "Corynebacteriales"),
    names(taxalist))],
    values = c(Streptomycetales = "limegreen", 
               Propionibacteriales = "steelblue1",
               Corynebacteriales = "goldenrod1" ))


## Add categorical results heatmap to base tree
figure_S4b <- gheatmap(q,
                       results_hm,
                       offset = .035,
                       width = 0.42,
                       color = "white",
                       colnames_angle = 55,
                       font.size = 1.6,
                       colnames_offset_y = 152,
                       colnames_offset_x = -0.01
) +
  scale_fill_manual(
    values = c("0" = "grey95",
               "1" = brewer.pal(11, "RdGy")[5],
               "2" = brewer.pal(11, "RdGy")[4],
               "3" = brewer.pal(11, "RdGy")[3],
               "4" = brewer.pal(11, "RdGy")[2]),
    labels = c("Inactive",
               "Low",
               "Medium",
               "High",
               "Outstanding"),
    na.value = "white",
    name = "Activity")


## Clean up
rm(p, q, name_order)
detach(list_tree_small_fil)

print(figure_S4b)

```

## Assemble *Figure S4*

```{r, create_figure_S4}

## Extract each legend individually, suppressing the other guide when doing so
legend_color1 <- cowplot::get_legend(figure_S4a + theme(legend.position = "right") + guides(fill = "none"))
legend_fill   <- cowplot::get_legend(figure_S4a + theme(legend.position = "right") + guides(color = "none"))
legend_color2 <- cowplot::get_legend(figure_S4b + theme(legend.position = "right") + guides(fill = "none"))

## Remove all legends from the tree plots themselves
figure_S4a <- figure_S4a + theme(legend.position = "none")
figure_S4b <- figure_S4b + theme(legend.position = "none")

## Stack the three legends in the desired order
legends_stacked <- wrap_plots(
  legend_color1, legend_fill, legend_color2,
  ncol = 1
)

figure_preS4 <- (wrap_elements(full = figure_S4a) / wrap_elements(full = figure_S4b)) +
  plot_layout(heights = c(1, 1.1))

figure_S4 <- (figure_preS4 | legends_stacked) +
  plot_layout(widths = c(1, 0.35))


rm(figure_preS4)

figure_S4

``` 


```{r save_figure_S4, include = FALSE}

ggsave("Figures/Supp_Figure_S4.pdf",
       figure_S4,
       height = 20,
       width = 17,
       units = "cm")

ggsave("Figures/Supp_Figure_S4.png",
       figure_S4,
       height = 20,
       width = 17,
       units = "cm",
       dpi = 600)

```

# End

```{r sessionInfo}

sessionInfo()

```


